Symbolic regression research is increasingly constrained by engineering fragmentation rather than a lack of methods alone. Modern symbolic regression systems differ substantially in repository structure, dependency stacks, execution assumptions, expression formats, and benchmark conventions, making it difficult to reproduce, compare, and extend methods under a unified protocol. We present \textbf{SR-Workbench}, a unified research platform for symbolic regression that standardizes both heterogeneous toolchains and heterogeneous datasets. On the method side, SR-Workbench wraps representative symbolic regression systems under a common interface with isolated runtime environments and reproducible execution paths. On the data side, it canonicalizes multiple symbolic regression dataset families into a lightweight schema with training, validation, in-distribution test, out-of-distribution test, metadata, and optional ground-truth formulas.

Built on top of this substrate, we introduce a modular skill layer for repository onboarding, dataset homogenization, protocol compilation, expression normalization, result auditing, and bounded improvement. These skills support a lightweight, auditable evolution loop that can reproduce baselines, validate execution, detect protocol violations, and propose constrained follow-up modifications without changing benchmark semantics. Rather than proposing a new symbolic regression model, SR-Workbench targets the research infrastructure bottleneck that slows down fast and reliable progress in symbolic regression. We position the current system as a practical foundation for reproducible symbolic regression experimentation, OOD-aware evaluation, and future automated method improvement.
