\section{Conclusion}

We presented SR-Workbench, a unified symbolic regression research platform that standardizes heterogeneous methods, heterogeneous datasets, and recurring research operations under a common substrate. Rather than introducing yet another symbolic regression model, our focus is on the infrastructure bottleneck that currently slows down symbolic regression research: repository fragmentation, environment incompatibility, inconsistent dataset conventions, and weak protocol control. SR-Workbench addresses these issues through a unified method interface, a canonical split-aware dataset schema, a modular skill layer, and a lightweight audited evolution loop.

The current version already demonstrates three practical capabilities. First, it integrates a diverse set of symbolic regression methods behind a common execution interface while respecting environment isolation. Second, it supports schema-level dataset validation, including formula-grounded verification when analytical ground truth is available. Third, it executes pilot end-to-end runs that expose both in-distribution and out-of-distribution behavior under a shared result format. These properties make the platform immediately useful as a reproducible experimentation substrate, even before the full automated loop is scaled up.

At the same time, we deliberately treat the current paper as a first systems release rather than a final benchmarking claim. Several components remain to be strengthened in the next iteration: larger-scale frozen-budget comparisons across all supported methods, more explicit expression normalization and complexity analysis, richer protocol compilation, and a more mature reviewer-guided improvement scheduler. We also plan to expand the current dataset substrate beyond the present pilot slice and further stress-test the platform on broader benchmark families such as SRBench, LLM-SRBench, and SRSD under a single audited protocol.

We believe the broader message is simple. In symbolic regression, progress is often limited not only by search algorithms, but also by the absence of reliable research infrastructure. By making method onboarding, dataset canonicalization, validation, and iterative execution explicit and reusable, SR-Workbench provides a concrete step toward faster, more reproducible, and more extensible symbolic regression research.
